---
title: "Data Import"
output: github_document
---


```{r}
library(tidyverse)
library(readxl)
library(haven)
```

let's import a dataset.

```{r}
litters_df=
  read_csv("data_import_examples/FAS_litters.csv")
```

```{r}
names(litters_df)
```

upload the names in 'litters_df'.

```{r}
litters_df = 
  janitor::clean_names(litters_df)
```

look at the data

```{r}
litters_df
```

sometimes skimming data is neat?
```{r}
skimr::skim(litters_df)
```

#Fix the missingness
```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""), skip = 5)
```

#import PUPS data
```{r}
pups_df = 
  read_csv(
    "data_import_examples/FAS_pups.csv",
    na = c("NA", ".", ""), 
    skip = 3,
    )
pups_df = 
  janitor::clean_names(pups_df)
```


# About excel
CSVs are really great but sometimes you get an excel file.

```{r}
mlb_df = 
  read_excel("data_import_examples/mlb11.xlsx")
```

Import LotR word counts
```{r}
fotr_df = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6")
```

# SAS
```{r}
Pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

Pulse_df = 
  janitor::clean_names(Pulse_df)
```


#why hate readcsv
```{r, eval=FALSE}
litters_df_base =
  read.csv("data_import_examples/FAS_litters.csv")
```

#what about data exporting
```{r}
write_csv(fotr_df, "data_import_examples/fort_df.csv")
```

#pivot_longer
```{r}
pulse_df = 
  haven::read_sas("data_import_examples/public_pulse_data.sas7bdat") |>
  janitor::clean_names()|>
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_" #delete the name of...
  ) |>
  mutate(visit = replace(visit, visit == "bl", "00m")) |>
  #relocate(id, visit)#把id visit 提前
  relocate(id, age)


pulse_df
```

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))|>
  janitor::clean_names()|>
  pivot_longer(
    cols = gd0_weight:gd18_weight,
    names_to = "gd_time",
    values_to = "weight",
  )|>
  mutate(gd_time = case_match(
    gd_time,
    "gd0_weight" ~ 0,
    "gd18_weight" ~ 18
  ))
```


#pivot_wider
let's make up an analysis result table
```{r}
analysis_df = 
  tibble(group = c("treatment", "treatment", "control","control"),
         time = c("pre", "post", "pre", "post"),
         mean = c(4, 10, 4.2, 5))
         
```

```{r}
analysis_df|>
  pivot_wider(
    names_from = time,
    values_from = mean
  )|>
  knitr::kable()
```

#bind table
```{r}
fellowship_ring = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_ring, two_towers, return_king)|>
  janitor::clean_names()|>
  pivot_longer(
    cols = female:male,
    names_to = "sex",
    values_to = "words"
  )|>
  relocate(movie)


```


#join FAS datasets
import "litters"
```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))|>
  janitor::clean_names()|>
  mutate(wt_gain = gd18_weight - gd0_weight)|>
  separate(
    group, into = c("dose", "day_of _treatment"), sep = 3
  )
```

import "pups"
```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv",na = c("NA", ".", ""), skip = 3)|>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )
```

join the dataset
```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")
  #relocate(litter_number, dose, day_of_treatment)
```



